\begin{frame}{GDP}
    \emph{Gross Domestic Product} measures the value of all goods and
    services produced in a country in a fixed period.
    \[GDP = consumption + investment + (exports - imports)\]
    Consumption and investment are further divided between private
    interests and government activities.

    \emph{Real GDP} corrects for inflation (measured independently) to make
    GDP comparable from year to year.
\end{frame}

\begin{frame}{Economic Indicators}
    \begin{description}
        \item [Treasury Interest Rate]
            Specifically, for 3-month treasury bills.  Set by the treasury to
            influence economic activity.
        \item [Total Capacity Utilization]
            The value of industrial goods produced compared to the available
            industrial capacity.
        \item [Industrial Production Index - IPI]
            Output from the manufacturing, mining, electric and gas industries. 
            Each individual index is calculated using the Fischer index formula.
        \item [Purchasing Manager Index]
            Index of industrial output and stock purchases gathered by survey.
        \item [Real Retail Sales]
            All goods sold by retailers based on a 
            sample of retail stores of different types and sizes.
    \end{description}
\end{frame}

\begin{frame}{Economic Indicators}
    \begin{description}
        \item [Civilian Labor Force Participation]
            The fraction of the eligible civilian labor force that is
            employed or seeking employment.
        \item [Real Personal Income]
            Average individual income, adjusted for inflation.
        \item [Non-farm employment]
            Total employment not counting the government, non-profits, and farms.
        \item [Outstanding Mortgage Debt]
            The total private mortgage debt.
        \item [Consumer Debt]
            Debts incurred on purchase of goods that are consumable and/or 
            do not appreciate.
    \end{description}
\end{frame}

\begin{frame}{Economic Indicators}
    \begin{description}
        \item [Private Consumption Expenditure]
            Total spending on goods and services targeted to and consumed by individuals.
        \item [Government Consumption Expenditure]
            Government spending which is not a transfer payment of money collected 
            in taxation from one group in society to another.
        \item [Gross Private Domestic Investment]
            Private sector investments including Capital, Inventory, and Residential assets.
    \end{description}
\end{frame}

\begin{frame}{Data Sources}
    Data drawn from the Federal Reserve Bank of St. Louis:
    
    \begin{center} \url{research.stlouisfed.org/fred2} \end{center}

    All data is quarterly (3-month) average and already seasonally adjusted.
\end{frame}